---
title: "Task5"
author: "Shantanu Patne"
date: "9/12/2020"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 16, fig.height = 12, message = FALSE, warning = FALSE)
```
# Business Analytics

```{r import, echo=FALSE, results='hide'}
library(ggplot2)
library(gridExtra)
library(dplyr)
```

## Data Preprocessing

### Read data

```{r read}
store <- read.csv("SampleSuperstore.csv")
head(store)
```


### Check for NAs and duplicates

```{r nrows}
nrow(store)
```

```{r check duplicates}
nrow(distinct(store))
```

Thus, we can see duplicates are present.

```{r check NA}
nrow(which(is.na(store)))
```

There aren't any missing values. Thus we need only to drop duplicates.

```{r drop}
store <- distinct(store)
nrow(store)
```

## State-wise analysis

```{r state}
# we need to see the overall Sales and Profits in each State
state_data <- store %>% 
  select(State, Sales, Profit, Discount) %>% 
    group_by(State) %>% 
      summarise(Sale = sum(Sales), Profits = sum(Profit), Discounts = sum(Discount), .groups = "keep")

state_data
```

```{r plot-st1}
state_sale <- ggplot(state_data, aes(x = reorder(State,Sale), y = Sale/1000)) + geom_col(fill = "darkblue")
state_sale <- state_sale + theme(axis.text.x = element_text(angle = 90, size = 15, vjust = 0.1), 
                                     axis.title = element_text(size = 20),
                                     axis.text.y = element_text(size = 15),
                                     title = element_text(size = 25))
state_sale <- state_sale + labs(x = "States", y = "Total Sales (in thousands)", title = "Total Sales per State")
state_sale
```

```{r plot-st2}
state_profit <- ggplot(state_data, aes(x = reorder(State, Profits), y = Profits/1000)) + geom_col(fill = "darkgreen")
state_profit <- state_profit + theme(axis.text.x = element_text(angle = 90, size = 15, vjust = 0.1), 
                                     axis.title = element_text(size = 20),
                                     axis.text.y = element_text(size = 15),
                                     title = element_text(size = 25))
state_profit <- state_profit + labs(x = "States", y = "Total Profits (in thousands)", title = "Total Profit per State")
state_profit
```

```{r plot-st3}
state_disc <- ggplot(state_data, aes(x = reorder(State, Discounts), y = Discounts)) + geom_col(fill = "darkred")
state_disc <- state_disc + theme(axis.text.x = element_text(angle = 90, size = 15, vjust = 0.1), 
                                     axis.title = element_text(size = 20),
                                     axis.text.y = element_text(size = 15),
                                     title = element_text(size = 25))
state_disc <- state_disc + labs(x = "States", y = "Total Discount", title = "Total Discount per State")
state_disc
```

#### From these three plots, we can see that the States providing the greatest discounts are the ones facing the greatest losses, i.e. Texas.

## Category-wise analysis

### Extract necessary data
```{r category}
cat_data <- store %>% 
  select(Category, Sales, Profit, Discount, Quantity,) %>% 
    group_by(Category) %>% 
      summarise(Sale = sum(Sales), Profits = sum(Profit), 
                Discounts = sum(Discount), Quantity = sum(Quantity), .groups = "keep")

cat_data
```
### Create a sales donut
```{r plot-cat1}
# Sales across categories
sdata <- cat_data %>% select(Category, Sale)
sdata$fraction <- sdata$Sale / sum(sdata$Sale)

# Compute the cumulative percentages (top of each rectangle)
sdata$ymax <- cumsum(sdata$fraction)

# Compute the bottom of each rectangle
sdata$ymin <- c(0, head(sdata$ymax, n=-1))

# Compute label position
sdata$labelPosition <- (sdata$ymax + sdata$ymin) / 2

# Compute a good label
sdata$label <- sprintf("%0.2f%%", sdata$fraction*100)

# Make the plot
p1 <- ggplot(sdata, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Category)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=Category), size=6) +
  scale_fill_brewer(palette="Dark2") +
  scale_color_brewer(palette="Dark2") +
  coord_polar(theta="y") +
  xlim(c(0, 4)) +
  theme_void() + labs(title = "Sales across Category") +
  theme(title = element_text(size = 25), legend.text = element_text(size = 15))

```

### Create a profits donut
```{r plot-cat2}
# Profit across categories
pdata <- cat_data %>% select(Category, Profits)
pdata$fraction <- pdata$Profits / sum(pdata$Profits)

# Compute the cumulative percentages (top of each rectangle)
pdata$ymax <- cumsum(pdata$fraction)

# Compute the bottom of each rectangle
pdata$ymin <- c(0, head(pdata$ymax, n=-1))

# Compute label position
pdata$labelPosition <- (pdata$ymax + pdata$ymin) / 2

# Compute a good label
pdata$label <- sprintf("%0.2f%%", pdata$fraction*100)

# Make the plot
p2 <- ggplot(pdata, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Category)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=Category), size=6) +
  scale_fill_brewer(palette="Dark2") +
  scale_color_brewer(palette="Dark2") +
  coord_polar(theta="y") +
  xlim(c(0, 4)) +
  theme_void() + labs(title = "Profit across Category") +
  theme(title = element_text(size = 25), legend.text = element_text(size = 15))

```

### Create a quantity donut
```{r plot-cat3}
# quantity across categories
qdata <- cat_data %>% select(Category, Quantity)
qdata$fraction <- qdata$Quantity / sum(qdata$Quantity)

# Compute the cumulative percentages (top of each rectangle)
qdata$ymax <- cumsum(qdata$fraction)

# Compute the bottom of each rectangle
qdata$ymin <- c(0, head(qdata$ymax, n=-1))

# Compute label position
qdata$labelPosition <- (qdata$ymax + qdata$ymin) / 2

# Compute a good label
qdata$label <- sprintf("%0.2f%%", qdata$fraction*100)

# Make the plot
p3 <- ggplot(qdata, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Category)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=Category), size=6) +
  scale_fill_brewer(palette="Dark2") +
  scale_color_brewer(palette="Dark2") +
  coord_polar(theta="y") +
  xlim(c(0, 4)) +
  theme_void() + labs(title = "Quantity across Category") +
  theme(title = element_text(size = 25), legend.text = element_text(size = 15))
```

### Create a discount donut
```{r plot-cat4}
# Discount across categories
ddata <- cat_data %>% select(Category, Discounts)
ddata$fraction <- ddata$Discounts / sum(ddata$Discounts)

# Compute the cumulative percentages (top of each rectangle)
ddata$ymax <- cumsum(ddata$fraction)

# Compute the bottom of each rectangle
ddata$ymin <- c(0, head(ddata$ymax, n=-1))

# Compute label position
ddata$labelPosition <- (ddata$ymax + ddata$ymin) / 2

# Compute a good label
ddata$label <- sprintf("%0.2f%%",ddata$fraction*100)

# Make the plot
p4 <- ggplot(ddata, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Category)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color = Category),size=6) +
  scale_fill_brewer(palette="Dark2") +
  scale_color_brewer(palette="Dark2") +
  coord_polar(theta="y") +
  xlim(c(0, 4)) +
  theme_void() + labs(title = "Discount across Category") +
  theme(title = element_text(size = 25), legend.text = element_text(size = 15))

```

### Display all donut charts
```{r}
grid.arrange(p1, p2, p3, p4, nrow=2)
```


#### From these plot we can incur that Technology category provides the highest profits while selling the lowest quantities.

## Subcategory analysis

```{r subcatprof}
sub <- store %>% select(Category, Sub.Category, Profit) %>%
  group_by(Category, Sub.Category) %>% summarise(Profit = sum(Profit), .groups = "keep")
subcat <- ggplot(sub, aes(x = reorder(Sub.Category, Profit), y = Profit/1000, fill = Category)) + geom_col()
subcat <- subcat + theme(axis.text.x = element_text(angle = 90, size = 15, vjust = 0.1), 
                         axis.title = element_text(size = 20),
                         axis.text.y = element_text(size = 15),
                         title = element_text(size = 25),
                         legend.text = element_text(size = 15)) + scale_fill_brewer(palette = "Dark2")
subcat<- subcat + labs(x = "Subcategory", y = "Profit (in thousand)", title = "Profit per Subcategory") 
subcat
```
#### From this we can see which subcategories show the highest profits and the greatest losses.

```{r subcatsale}
sub2 <- store %>% select(Category, Sub.Category, Sales) %>%
  group_by(Category, Sub.Category) %>% summarise(Sales = sum(Sales), .groups = "keep")
subcat2 <- ggplot(sub2, aes(x = reorder(Sub.Category, Sales), y = Sales/1000, fill = Category)) + geom_col()
subcat2 <- subcat2 + theme(axis.text.x = element_text(angle = 90, size = 15, vjust = 0.1), 
                         axis.title = element_text(size = 20),
                         axis.text.y = element_text(size = 15),
                         title = element_text(size = 25),
                         legend.text = element_text(size = 15)) + scale_fill_brewer(palette = "Dark2")
subcat2<- subcat2 + labs(x = "Subcategory", y = "Sales (in thousand)", title = "Sales per Subcategory") 
subcat2
```
#### From this we can see which subcategories show the greatest sales.

```{r subcatdisc}
sub3 <- store %>% select(Category, Sub.Category, Quantity) %>%
  group_by(Category, Sub.Category) %>% summarise(Q = sum(Quantity), .groups = "keep")
subcat3 <- ggplot(sub3, aes(x = reorder(Sub.Category, Q), y = Q, fill = Category)) + geom_col()
subcat3 <- subcat3 + theme(axis.text.x = element_text(angle = 90, size = 15, vjust = 0.1), 
                         axis.title = element_text(size = 20),
                         axis.text.y = element_text(size = 15),
                         title = element_text(size = 25),
                         legend.text = element_text(size = 15)) + scale_fill_brewer(palette = "Dark2")
subcat3<- subcat3 + labs(x = "Subcategory", y = "Quantity", title = "Amount per Subcategory") 
subcat3
```

#### From these plots we can also see that the Copiers, which sold the least quantities brought in the greatest profit.

## Segment based analysis

### Separaing necessary data
```{r seg1}
seg <- store %>% 
  select(Segment, Sales, Profit, Category) %>%
    group_by(Segment, Category) %>%
    summarise(Sales = sum(Sales), Profit = sum(Profit), .groups = "keep")
seg
```

### Plotting Profit vs Segment
```{r segprof}
seg_prof <- ggplot(seg, aes(x = reorder(Segment, Profit), y = Profit, fill=Category)) + geom_col(position = "dodge") 
seg_prof <- seg_prof + theme(axis.text.x = element_text(size = 15, vjust = 0.1), 
                         axis.title = element_text(size = 20),
                         axis.text.y = element_text(size = 15),
                         title = element_text(size = 25),
                         legend.text = element_text(size = 15)) + scale_fill_brewer(palette = "Dark2")
seg_prof <- seg_prof + labs(x = "Segment", y = "Profit", title = "Profit per segment", subtitle = "Separated by Category")
seg_prof
```

### Plotting Sales vs Segment
```{r segsale}
seg_sale <- ggplot(seg, aes(x = reorder(Segment, Sales), y = Sales, fill=Category)) + geom_col(position = "dodge") 
seg_sale <- seg_sale + theme(axis.text.x = element_text(size = 15, vjust = 0.1), 
                         axis.title = element_text(size = 20),
                         axis.text.y = element_text(size = 15),
                         title = element_text(size = 25),
                         legend.text = element_text(size = 15)) + scale_fill_brewer(palette = "Dark2")
seg_sale <- seg_sale + labs(x = "Segment", y = "Sales", title = "Sales per segment", subtitle = "Separated by Category")
seg_sale
```

From this graph we can analyze which Segment to target for any given Category.

























